validateHOT - an R package for holdout task validation and market simulations

### 2024-01-12 paper/paper.Rmd
try compiling - 240104
updated figure and turf() manual - 240105
updated intro of paper - 240112

---
title: 'validateHOT - an R package for holdout task validation and market simulation'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 02 January
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
 - \usepackage{fancyhdr}
---

```{=tex}
\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```

```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way for testing validity is by including so-called validation or holdout tasks [e.g., @rao2014; @Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities in hierarchical Bayes estimation. Practitioners often do not include them [@yang2018], which is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

validateHOT combines both validation and market simulation in one package and has three key advantages, it (1) helps to opt for the best model, (2) runs relevant market simulations that help to find the right product combinations or assortments, and finally, (3) is an open source tool including functions that are usually implemented in property commercial, and therefore, remain a black-box for researchers and practitioners.

@@

Other packages provide functions to calculate validation metrics, however, these are not specified for individual raw logit coefficients which are usually the output when running random parameter logit / hierarchical Bayes models. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones. However, no functions for validation are included and moreover, conjoint [@conjoint] focuses on classical conjoint analysis, and thus is limited when applying more common CBC methods, for example, (A)CBC. support.BWS [@support.BWS] only covers best-worst scaling case 1 (also known as MaxDiff) and only provides market simulations based on conditional logit rule. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. A comparison of validateHOT's functions with current R packages is provided in \autoref{comparison}. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```

validateHOT is introduced with data estimated with Lighthouse Studio using effects-coding for creating the design matrix. It, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

@@

@@

>>>>

---
title: 'validateHOT - an R package for holdout task validation and market simulations'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Benjamin Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1, 2
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 02 January
year: 2024
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
header-includes: 
 - \usepackage{float}
 - \usepackage{fancyhdr}
---

```{=tex}
\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```
```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

Preference measurement techniques' (e.g., (A)CBC or MaxDiff) aim is to predict behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and predicts outside tasks (i.e., the model has external validity) well.[^1] The easiest way for testing validity is by including so-called validation or holdout tasks [e.g., @rao2014; @Orme2015], which are tasks that are fixed (i.e., same across participants) and are usually not used for estimating the part-worth utilities (raw logit utilities) in hierarchical Bayes estimation. Despite the importance of validation/holdout tasks, practitioners do not always include them [@yang2018]. This is unsatisfactory given the fact that the model is used to estimate market shares which poses the basis for relevant marketing decisions.

@@

validateHOT combines both validation and market simulation in one package and has three key advantages, it (1) helps to opt for the best model and (2) runs relevant market simulations that help to find the right product combinations or assortments, and finally, (3) is an open source tool which helps especially researchers to report accompanied scripts for their research papers.

@@

Other packages provide functions to calculate validation metrics, however, these are not specified for individual raw logit coefficients which are usually the output when running random parameter logit / hierarchical Bayes models. Metrics [@Metrics], for example, provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix. However, to get the output of, for example, Sawtooth Software or ChoiceModelR [@ChoiceModelR] into the right format, the user needs some data wrangling. The package conjoint [@conjoint] provides functions that are most similar to validateHOT's ones. However, no functions for validation are included and moreover, conjoint [@conjoint] focuses on classical conjoint analysis, and thus is limited when applying more common conjoint methods, for example, (A)CBC. support.BWS [@support.BWS] only covers best-worst scaling case 1 (also known as MaxDiff) and only provides market simulations based on conditional logit rule. logitr [@logitr] provides market simulations tools, however, no validation metrics such as mean hit probability [@voleti2017] or hit rate [@netzer2011]. A comparison of validateHOT's functions with current R packages is provided in \autoref{comparison}. To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations (especially TURF) is missing.

```{=tex}
\begin{figure}[h]
  \includegraphics{figures/FunctionComparison.png}
  \caption{Comparison of validateHOT's function to existing R packages}
  \label{comparison}
\end{figure}
```
validateHOT is introduced with data estimated with Lighthouse Studio using effects-coding for creating the design matrix. It, however, can easily be used with data estimated with ChoiceModelR [@ChoiceModelR], bayesm [@bayesm], or STAN [@rstan], if used with similar settings (ChoiceModelR, for example, automatically implements effects-coding).

@@

\newpage

@@

# Acknowledgments

We would like to thank [Sawtooth Software](https://sawtoothsoftware.com/) for their great transparent documentation.
